---
title: "&nbsp;"
output:
  html_document:
    theme:
      version: 4
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## problem 6 and 7

### Import data
First, Set the enviroment and import data.

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
rm(list=ls())
library(haven) #library for importing data set
library(psych) #library for trace matrix
library(lmtest) #library for HC3
library(sandwich) #library for HC3
library(Statamarkdown)
cps09mar <- read_dta("cps09mar.dta")
```

```{r echo=FALSE}
knitr::kable(cps09mar[1:10,1:12], format="markdown")
```

*Note for region and martial status*

####   region:

-   1: Northeast

-   2: Midwest

-   3: South

-   4: West

####   Marital status:

-   1: Married - civilian spouse present

-   2: Married - Armed Forces spouse present

-   3: Married - spouse absent (except separated)

-   4: Widowed

-   5: Divorced

-   6: Separated

-   7: Never married


### Data Manipulation

Slice the data for White male Hispanics,
```{r}
cps <- cps09mar[cps09mar$hisp == 1 & cps09mar$female == 0 & cps09mar$race == 1,] 
```

And create matrix for variables.

```{r}
con <- matrix(1,nrow = nrow(cps),ncol = 1)
edu <- matrix(0,nrow = nrow(cps),ncol = 1)
edu[,] <- cps$education

exp <- matrix(0,nrow = nrow(cps),ncol = 1)
exp[,] <- cps$age - cps$education - 6

expsq <- exp^2

wage <- matrix(0,nrow = nrow(cps),ncol = 1)
wage[,] <- cps$earnings / (cps$hours * cps$week)
lwage <- log(wage)

#region dummy variables

dne <- matrix(0,nrow = nrow(cps),ncol = 1)
ds <- matrix(0,nrow = nrow(cps),ncol = 1)
dw <- matrix(0,nrow = nrow(cps),ncol = 1)
dne[cps$region == 1,] <-  1
ds[cps$region == 3,] <-  1
dw[cps$region == 4,] <-  1

#Martial dummy variables

dmar <- matrix(0,nrow = nrow(cps),ncol = 1)
dwid <- matrix(0,nrow = nrow(cps),ncol = 1)
dsep <- matrix(0,nrow = nrow(cps),ncol = 1)

dmar[cps$marital == 1 | cps$marital == 2 | cps$marital == 3,] <-  1
dwid[cps$marital == 4 | cps$marital == 5,] <-  1
dsep[cps$marital == 6,] <-  1


X <- cbind(con,edu,exp,expsq,dne,ds,dw,dmar,dwid,dsep)
Y <- lwage
```

*Note that cps is subsample of white male hispanics.*

### Estimate a log wage regression(3.26)

Now I can estimate Beta and Error.

```{r}

beta_hat <- solve((t(X) %*% X)) %*% (t(X) %*% Y)
e_hat <- Y - X %*% beta_hat
row.names(beta_hat) <- c('con','edu','exp','expsq',
                         'dne','ds','dw','dmar','dwid','dsep')
colnames(beta_hat) <- c('coefficient')
colnames(e_hat) <- c('error')

```

```{r include=FALSE}
library(kableExtra)
```
<div style='page-break-before:always'></div>

```{r echo=FALSE, message=FALSE, warning=FALSE}

knitr::kable(beta_hat,caption = 'Beta estimation',table.attr = "style='width:30%;'") %>% 
  kable_styling(full_width = TRUE, position = 'float_left')

knitr::kable(e_hat[1:10,],col.names = 'error',caption =
               'Error, First 10 elements',table.attr = "style='width:20%;'") %>% 
  kable_styling(full_width = TRUE, position = 'left')
```

Now, I'll estimate same thing with Stata.


```{stata collectcode = TRUE}
clear
use "/Users/hjr/Desktop/econometrics_pset1/cps09mar.dta"
keep if (hisp == 1 & female == 0 & race == 1)
gen lwage = log(earnings/(hours*week))
gen exp = age - education - 6
gen expsq = exp^2
gen dne = region == 1
gen ds = region == 3
gen dw = region == 4
gen dmar = marital == 1 | marital ==2 | marital == 3
gen dwid = marital == 4 | marital == 5
gen dsep = marital == 6
regress lwage education exp expsq dne ds dw dmar dwid dsep
```

### Calculate standard errors using the HC3 method(4.25)

Set the variable for convenience.
```{r}
XX <- solve((t(X) %*% X)) #XX is variable name it means inverse matrix of (X'X)
leverage <- rowSums(X*(X %*% XX))
```

Calculate S.E using the HC3 method.
```{r}
u3 <- X * ((e_hat/(1-leverage))%*% matrix(1,1,ncol(X)))
v3 <- XX %*% (t(u3) %*% u3) %*% XX
s3 <- sqrt(diag(v3))
```

```{r echo=FALSE}
se <- data.frame(s3,row.names = c('con','edu','exp','expsq',
                            'dne','ds','dw','dmar','dwid','dsep'))
se <- t(se)
row.names(se) <- 'S.E.'
```

```{r echo = F}
knitr::kable(se,caption = 'S.E.(HC3)') %>% 
  kable_styling(full_width = TRUE, position = 'center', font_size = 12)
```

Then calculate with Stata.
```{stata}
regress lwage education exp expsq dne ds dw dmar dwid dsep, vce(hc3)
```

Yield same results.